---
title: "Experiment test"
output: 
  revealjs::revealjs_presentation:
    transition: none
    
date: "2024-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(lubridate)
library(ggplot2)
library(data.table)
library(metR)
library(rcdo)

cdo_options_set("-L -s")

source(here::here("R/functions.R"))
source(here::here("R/datasets.R"))
source(here::here("R/ggplot.R"))

zero_breaks <- AnchorBreaks(0, exclude = 0)

```


```{r}
forecast <- function(variable, experiments) {
  vapply(experiments, function(experiment) {
    file <- glue::glue("/home/565/ec0044/cylc-run/202405_{experiment}/share/data/20240501T0000Z/forecast/20240501/e01/d*_{variable}_*_e01.nc")  |> 
      Sys.glob()
  }, character(1))
}
```


```{r}
ic <- function(part, experiments) {
  vapply(experiments, function(experiment) {
    file <- glue::glue("/home/565/ec0044/cylc-run/202405_{experiment}/share/data/20240501T0000Z/gc2-IC-bom/{part}/*") |> 
      Sys.glob()
  }, character(1))
  
}
```

```{r}
experiments <- c("control", 1990)
initial <- as.Date("2024-05-01")
```

```{r}
cdo_remap1 <- function(file) {
  file |> 
    cdo_setmisstonn(neighbors = 2) |> 
    cdo_remapavg(grid = "r360x180")
}
```

```{r}
climatology_cdr <- CDR() |> 
  climatology() |> 
  ReadNetCDF(c(clim = "aice")) |> 
  setnames("time", "time2")
```

## POC for sea ice experiments



## Setup

* Initial conditions: 1st May 2024
* Control run: regular initial conditions. 
* These are S2 initial conditions. 
* Pre-DA initial conditions -> need to check up with Griff/Lynn on what does that exactly mean.

* Experiment run: sea ice initial conditions set up to 1990 (random round number)


## Sea Ice concentration anomalies

```{r}
ice_ic <- forecast("aice", experiments) |> 
  lapply(\(x) {
    x |> 
      cdo_seldate(startdate = as.character(initial + 1)) |> 
      cdo_selname("aice") |> 
      remap_cdr() |> 
      cdo_execute() |>
      ReadNetCDF(vars = c("aice"))
  }) |> 
  rbindlist(idcol = "experiment") |> 
  _[, time2 := update(time, year = 2000)]


ice_ic <- climatology_cdr[ice_ic, on = .NATURAL]
```

```{r}
ice_ic |> 
  ggplot(aes(xgrid, ygrid)) +
  geom_contour_fill(aes(z = aice - clim, fill = after_stat(level)),
                    breaks = zero_breaks) +
  scale_fill_divergent_discretised(NULL, 
                                   low = scales::muted("red"),
                                   high = scales::muted("blue")) +
  geomcoord_antarctica +
  geom_antarctica_fill +
  facet_wrap(~ experiment)  +
  wide_legend +
  labs(subtitle = "Anomaly wrt NOAA CDR climatology")
```       


## Control - experiment

```{r}
ice_ic |> 
  copy() |> 
  _[, let(control = aice[experiment == "control"]), 
    by = .(ygrid, xgrid)] |> 
  _[experiment != "control"] |> 
  ggplot(aes(xgrid, ygrid)) +
  geom_contour_fill(aes(z = control - aice, fill = after_stat(level)),
                    breaks = zero_breaks) +
  scale_fill_divergent_discretised(NULL, 
                                   low = scales::muted("red"),
                                   high = scales::muted("blue")) +
  geomcoord_antarctica +
  geom_antarctica_fill +
  facet_wrap(~ experiment)  +
  wide_legend 
```

## Zonal mean zonal wind at 60ºS, 10 hPa. 

```{r}
forecast("ua", experiments) |> 
  lapply(\(x) {
    x |> 
      ReadNetCDF(c(u = "ua"), 
                 subset = list(z1_p_level = 10, lat = -60)) |> 
      _[, .(u = mean(u)), by = .(time)] 
  }) |> 
  rbindlist(idcol = "experiment") |> 
  ggplot(aes(time, u)) +
  geom_line(aes(color = experiment))
```

## Zonal mean zonal wind Hovmoller 


```{r}
forecast("ua", experiments) |> 
  lapply(\(x) {
    x |> 
      ReadNetCDF(c(u = "ua"), 
                 subset = list(lat = -60)) |> 
      _[, .(u = mean(u, na.rm = TRUE)), by = .(z1_p_level, time)] 
  }) |> 
  rbindlist(idcol = "experiment") |> 
  ggplot(aes(time, z1_p_level)) +
  geom_contour_fill(aes(z = u)) +
  scale_fill_viridis_c() +
  scale_y_level() +
  facet_wrap(~ experiment) 
```


## Control - experiment


```{r}
forecast("ua", experiments) |> 
  lapply(\(x) {
    x |> 
      ReadNetCDF(c(u = "ua"), 
                 subset = list(lat = -60)) |> 
      _[, .(u = mean(u, na.rm = TRUE)), by = .(z1_p_level, time)] 
  }) |> 
  rbindlist(idcol = "experiment") |> 
  copy() |> 
  _[, let(control = u[experiment == "control"]), 
    by = .(z1_p_level, time)] |> 
  _[experiment != "control"] |> 
  ggplot(aes(time, z1_p_level)) +
  geom_contour_fill(aes(z = control - u, fill = after_stat(level)),
                    breaks = zero_breaks) +
  scale_fill_divergent_discretised(NULL) +
  scale_y_level() +
  wide_legend
```

## Stratospheric temperature

```{r}
t <- forecast("ta", experiments) |> 
  lapply(\(x) {
    x |> 
      cdo_selname("ta") |> 
      cdo_seldate(startdate = "2024-06-01", enddate = "2024-06-15") |> 
      cdo_remapbil(grid = "r360x180") |>
      # cdo_timmean() |> 
      cdo_execute() |> 
      ReadNetCDF(c(t = "ta"), 
                 subset = list(z1_p_level = 10, lat = c(-90, -20))) 
  }) |> 
  rbindlist(idcol = "experiment") 

t |> 
  _[time == time[2]] |> 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = t - 273.15), proj = sic_projection, proj.latlon = FALSE) +
  scale_fill_viridis_c() +
  wide_legend +
  geom_antarctica_path +
  coord_sf(crs = sic_projection, lims_method = "box") +
  facet_grid(experiment ~ z1_p_level) +
  labs(substitle = "Averaged between 1 and 15 July") +
  

```



